Hexagonal binning algorithm

### 2024-05-19 paper.qmd
Added MNIST digit description, need to add the video

bibliography: bibliography.bib  
header-includes: | 
  \usepackage{amsmath}
  \usepackage{float}
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \usepackage{bm}
  \def\tightlist{}
  \usepackage{setspace}
  \newcommand\pD{$p\text{-}D$}
  \newcommand\kD{$k\text{-}D$}
  \newcommand\dD{$d\text{-}D$}
---

@@

```{r}
#| label: code-setup
set.seed(20240110)
source("nldr_code.R", local = TRUE)
```

@@

## Hexagonal binning to have regular hexagons
hb_obj_scurve <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = num_bins_x_scurve, 
  q = 0.1,
  r2 = r2)

@@

## To obtain the standardise counts within hexbins
counts_df <- hb_obj_scurve$std_cts
df_bin_centroids <- extract_hexbin_centroids(
  centroids_df = all_centroids_df, 
  counts_df = counts_df)

@@

@@

@@

pbmc_model <- fit_highd_model(
  training_data = training_data_pbmc,
  emb_df = umap_pbmc_scaled,
  bin1 = num_bins_x_pbmc,
  q = 0.1,
  r2 = r2_pbmc,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "PC_"
)

@@

pbmc_model <- fit_highd_model(
  training_data = training_data_pbmc,
  emb_df = tsne_pbmc_scaled,
  bin1 = num_bins_x_pbmc,
  q = 0.1,
  r2 = r2_pbmc,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "PC_"
)

@@

### digits: 1

- NLDR is used to illustrate different ways 1's are drawn
- Use our method to assess is it a reasonable representation
- Demonstrate that it is, except for the anomalies

@@

pacmap_plot_mnist <- pacmap_minst_scaled |> 
  ggplot(aes(x = PaCMAP1, 
             y = PaCMAP2)) + 
  geom_point(alpha=0.1) + 
  geom_text(data = position_df, 
            aes(x = PaCMAP1, y = PaCMAP2, label = position), 
            colour = "red", 
            size = 5)  
```

@@

img_sample <- pixels_gathered |> 
  ggplot(aes(x, y, fill = value)) +  
  geom_tile() +
  facet_wrap(~ factor(instance, levels = c(3710, 2391, 2385, 5030, 6475,
                         7679, 3568, 1312, 3311, 1097,
                         3552, 7853, 6489, 7689, 6690,
                        1380, 6057, 2347, 5946, 3355,
                        4175, 3997, 5378, 387, 1854,
                        614, 3079, 1762, 5239, 3723,
                        5748, 728, 7419, 7794, 6233,
                        33, 2485, 5998, 318, 1761,
                        5690, 165, 517, 6935, 2682,
                        4962, 2264, 5563, 6369, 559,
                        5188, 4849, 2666, 3448, 3055,
                        3120, 6869, 6345, 4470, 7147)), ncol = 5) +
  scale_fill_continuous_sequential(palette = "Grays") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(), 
        legend.position = "none")
```

<!--PaCMAP param: n_components=2, n_neighbors=10, init=random, MN_ratio=0.9, FP_ratio=2.0-->
```{r}
#| fig-cap: "(a) 2D layout from PaCMAP applied for the digit 1 of the MNIST dataset. Is this a best representation of the original data?, and (b) Images of the handwritten digit 1. The angle of the digit 1 varies along non-linear structure."
#| label: fig-pacmapauthor
#| fig-pos: H

pacmap_plot_mnist + img_sample +
  plot_annotation(tag_levels = 'a')
```

@@

mnist_model <- fit_highd_model(
  training_data = training_data_mnist,
  emb_df = pacmap_minst_scaled,
  bin1 = num_bins_x_mnist,
  q = 0.1,
  r2 = r2_mnist,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "PC"
)

@@

trimesh_removed_mnist <- vis_rmlg_mesh(
  distance_edges = distance_mnist, 
  benchmark_value = benchmark_mnist, 
  tr_coord_df = tr_from_to_df_mnist, 
  distance_col = "distance") +
  #xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("a", sc_ltr_pos)

@@

error_plot_mnist <- error_df |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2, 
             color = type, 
             group = ID)) +
  geom_point(alpha=0.5, 
             size = 0.1) +
  interior_annotation("b", sc_ltr_pos)
```

<!--bin1 = 37, bin2 = 26, b = 962, non_empty = 398-->
```{r}
#| echo: false
#| fig-cap: "(a) Model generated in the 2D space overlaid on PaCMAP data, and (b) high-D model error in model space."
#| label: fig-model-mnist
#| fig-pos: H

@@

@@

imge_error_sample_within <- pixels_gathered |> 
  ggplot(aes(x, y, fill = value)) +  
  geom_tile() +  
  facet_wrap(~ factor(instance, levels = c( 4699, 4785, 5173, 6671, 2186, 4636, 
                                            4882, 773, 2825, 6169, 6948, 5987, 
                                            1282, 7715, 622, 2989, 6912, 4370, 
                                            1227, 4042, 2444, 836, 1654)), 
             ncol = 5) +
  scale_fill_continuous_sequential(palette = "Grays") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none")
```

@@

imge_error_sample <- pixels_gathered |> 
  ggplot(aes(x, y, fill = value)) +  
  geom_tile() +  
  facet_wrap(~ instance, ncol = 2) +
  scale_fill_continuous_sequential(palette = "Grays") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none")
```

```{r}
#| echo: false
#| fig-cap: "(a) Images of handwritten digit 1 which occur large model error within the non-linear strcuture, and (b) Images of handwritten digit 1 which occur large model error outside the non-linear structure."
#| label: fig-anomalies
#| fig-width: 4
#| fig-height: 4
#| out-width: 100%
#| fig-pos: H

>>>>

bibliography: bibliography.bib  
header-includes: | 
  \usepackage{amsmath}
  \usepackage{float}
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \usepackage{bm}
  \def\tightlist{}
  \usepackage{setspace}
  \newcommand\pD{$p\text{-}D$}
  \newcommand\kD{$k\text{-}D$}
  \newcommand\dD{$d\text{-}D$}
  \newcommand\gD{$2\text{-}D$}
---

@@

```{r}
#| label: code-setup
set.seed(20240110)
#source("nldr_code.R", local = TRUE)
```

@@

## Hexagonal binning to have regular hexagons
hb_obj_scurve <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = num_bins_x_scurve, 
  r2 = r2)

@@

## To obtain the standardise counts within hexbins
counts_df <- hb_obj_scurve$std_cts
df_bin_centroids <- extract_hexbin_centroids(
  centroids_df = all_centroids_df, 
  counts_df = counts_df) |>
  filter(drop_empty == FALSE)

@@

<!--need to add what is meant by a long edge-->

@@

#### Fitted values

The prediction approach involves performing the K-nearest neighbors (KNN) algorithm for an unsupervised classification problem. First, the nearest high-D model point is identified for a given new high-D point. Then, the corresponding 2D centroid mapping for the identified high-D model point is determined. Finally, the coordinates of this 2D centroid are used as the predicted 2D embedding for the new high-D data point. This step is particularly valuable due to the limitations of some NLDR techniques, like tSNE, which don't provide a straightforward method for prediction. As a result, our approach offers a solution that capable of generating predicted 2D embedding regardless of the NLDR technique employed, effectively addressing this functional gap.

#### Error calculation

Residuals are essential for evaluating the accuracy of representing high-D points by the high-D mapping of 2D bin centroids. To measure this accuracy, an error metric is introduced, quantifying the sum of squared differences between the high-D data ($x_{ij}$) and the high-D mapping of the 2D bin centroid data ($C_{x_ij}$) across all observations and dimensions (see @eq-equation10).

$$
\text{Error} = \sum_{j = 1}^{n}\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2
$$ {#eq-equation10}

Here, $n$ represents the number of observations, $p$ represents the dimensions of high-D data, $x_{ij}$ is the high-D data, and $C_{x_ij}$ is the high-D mapping of the 2D bin centroid.

To assess how well our method captures and represents the underlying structure of the high-D data, Mean Squared Error (MSE) is used. When computing MSE, total model error (see @sec-summary) is divided by the number of observations to make it as a mean value (see @eq-equation9).

$$
\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{n}
$$ {#eq-equation9}

@@

pbmc_model <- fit_highd_model(
  training_data = training_data_pbmc,
  emb_df = umap_pbmc_scaled,
  bin1 = num_bins_x_pbmc,
  r2 = r2_pbmc,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "PC_"
)

@@

pbmc_model <- fit_highd_model(
  training_data = training_data_pbmc,
  emb_df = tsne_pbmc_scaled,
  bin1 = num_bins_x_pbmc,
  r2 = r2_pbmc,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "PC_"
)

@@

### digits: 1
<!--
- NLDR is used to illustrate different ways 1's are drawn
- Use our method to assess is it a reasonable representation
- Demonstrate that it is, except for the anomalies 
-->

The MNIST dataset consists of grayscale images of handwritten digits [@lecun2010]. @yingfan2021 used this dataset to demonstrate how PaCMAP preserves local structure. We selected the \gD{} embedding of PaCMAP for the handwritten digit 1 to assess whether this is a reasonable representation using our method. As shown in @fig-pacmap-author-img, the angle of the digit 1 images varies along the \gD{} structure.

@@

pacmap_plot_mnist <- pacmap_minst_scaled |> 
  ggplot(aes(x = PaCMAP1, 
             y = PaCMAP2)) + 
  geom_point(alpha=0.1) #+ 
  # geom_text(data = position_df, 
  #           aes(x = PaCMAP1, y = PaCMAP2, label = position), 
  #           colour = "red", 
  #           size = 5)  
```

<!--PaCMAP param: n_components=2, n_neighbors=10, init=random, MN_ratio=0.9, FP_ratio=2.0-->
```{r}
#| fig-cap: "$2\\text{-}D$ layout from PaCMAP applied for the digit 1 of the MNIST dataset. Is this the best representation of the digit 1? The parameter setting is n_components=2, n_neighbors=10, init=random, MN_ratio=0.9, FP_ratio=2.0."
#| label: fig-pacmap-author
#| fig-pos: H
#| out-height: 22%

pacmap_plot_mnist
```

@@

img_sample <- pixels_gathered |> 
  ggplot(aes(x, y, fill = value)) +  
  geom_tile() +
  facet_wrap(~ factor(instance, levels = c(3710, 2391, 2385, 5030, 6475,
                         7679, 3568, 1312, 3311, 1097,
                         3552, 7853, 6489, 7689, 6690,
                        1380, 6057, 2347, 5946, 3355,
                        4175, 3997, 5378, 387, 1854,
                        614, 3079, 1762, 5239, 3723,
                        5748, 728, 7419, 7794, 6233,
                        33, 2485, 5998, 318, 1761,
                        5690, 165, 517, 6935, 2682,
                        4962, 2264, 5563, 6369, 559,
                        5188, 4849, 2666, 3448, 3055,
                        3120, 6869, 6345, 4470, 7147)), nrow = 5) +
  coord_fixed() +
  scale_fill_continuous_sequential(palette = "Grays") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(), 
        legend.position = "none")
```

```{r}
#| fig-cap: "Images of the handwritten digit 1 are ordered from the bottom-right to the top-left of the $2\\text{-}D$ structure. The angle of the digit varies along this structure. Images at the bottom-right of the $2\\text{-}D$ layout show the digit 1 angled more to the right, while images at the top-left show the digit 1 angled more to the left. This demonstrates how the angle changes from right to left along the $2\\text{-}D$ structure."
#| label: fig-pacmap-author-img
#| fig-pos: H
#| out-height: 50%

img_sample
```

@@

mnist_model <- fit_highd_model(
  training_data = training_data_mnist,
  emb_df = pacmap_minst_scaled,
  bin1 = num_bins_x_mnist,
  r2 = r2_mnist,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "PC"
)

@@

sc_ltr_pos_mnist <- c(0.96, 0.96)

trimesh_removed_mnist <- vis_rmlg_mesh(
  distance_edges = distance_mnist, 
  benchmark_value = benchmark_mnist, 
  tr_coord_df = tr_from_to_df_mnist, 
  distance_col = "distance") +
  #xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("a", sc_ltr_pos_mnist)

@@

error_plot_mnist <- error_df |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2, 
             color = type, 
             group = ID)) +
  geom_point(alpha=0.5, 
             size = 0.1) +
  interior_annotation("b", sc_ltr_pos_mnist)
```

<!--bin1 = 37, bin2 = 26, b = 962, non_empty = 398-->
```{r}
#| echo: false
#| fig-cap: "(a) Model generated in $2\\text{-}D$, and (b) $p\\text{-}D$ model error in $2\\text{-}D$. The $2\\text{-}D$ model shows a non-linear continuous structure. Most low $p\\text{-}D$ model errors are distributed along the lower edge of the $2\\text{-}D$ structure, while most high p-D model errors are concentrated along the upper edge."
#| label: fig-model-mnist
#| fig-pos: H

@@

According to @fig-mnist1_sc1, the non-linear continuous structure observed in the \gD{} representation of PaCMAP (see @fig-pacmap-author) is also visible when visualizing the model overlaid on the data space. This indicates that PaCMAP accurately captures the structure of the \pD{} data. Additionally, the model shows a twisted pattern within the non-linear structure in \pD{} space (see @fig-mnist1_sc2), which is an additional pattern not visible in the \gD{} representation (see @fig-pacmap-author). Furthermore, as shown in @fig-mnist1_sc3, some long edges exist in the \pD{} space that are not recognized as long edges in the \gD{} representation. However, PaCMAP is a *good* \gD{} representation of MNIST digit 1 data.

<!--add langevitour screenshots and youtube animation link-->

::: {#fig-mnist1_sc layout-ncol="3" fig-pos="H"}
![](figures/mnist/sc_1.png){#fig-mnist1_sc1}

![](figures/mnist/sc_2.png){#fig-mnist1_sc2}

![](figures/mnist/sc_3.png){#fig-mnist1_sc3}

Screen shots of the **langevitour** of the MNIST digit 1 data set, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/zq2GM9qvUNA>).
:::
<!-- need to update the youtube recording-->

There are certain data points that exhibit high error rates due to their deviation from the usual \pD{} data structure, which makes them anomalies (see @fig-model-mnist (b)). These anomalies can be classified into two types: those that are anomalies within the non-linear structure and those that lie outside of it. The images associated with high model error points within the non-linear structure display different patterns of the digit 1, as shown in @fig-mnist-anomalies (a). However, when comparing these images to the ones found outside of the non-linear structure, it becomes evident that the latter display different patterns of the digit 1 (see @fig-mnist-anomalies (b)).

@@

imge_error_sample_within <- pixels_gathered |> 
  ggplot(aes(x, y, fill = value)) +  
  geom_tile() +  
  facet_wrap(~ factor(instance, levels = c( 4699, 4785, 5173, 6671, 2186, 4636, 
                                            4882, 773, 2825, 6169, 6948, 5987, 
                                            1282, 7715, 622, 2989, 6912, 4370, 
                                            1227, 4042, 2444, 836, 1654)), 
             ncol = 4) +
  coord_fixed() +
  scale_fill_continuous_sequential(palette = "Grays") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none")
```

@@

imge_error_sample <- pixels_gathered |> 
  ggplot(aes(x, y, fill = value)) +  
  geom_tile() +  
  facet_wrap(~ instance, ncol = 2) +
  coord_fixed() +
  scale_fill_continuous_sequential(palette = "Grays") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none") 
```

```{r}
#| echo: false
#| fig-cap: "Some images of handwritten digit 1 which occur high model error (a) within the non-linear strcuture, and (b) outside the non-linear structure. The images shows different patterns of digit 1."
#| label: fig-mnist-anomalies
#| out-height: 35%
#| fig-pos: H

### 2024-05-19 paper.tex
Added MNIST digit description, need to add the video

\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-1in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.7in}%
\addtolength{\topmargin}{-1in}%
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{float}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{bm}
\def\tightlist{}
\usepackage{setspace}
\newcommand\pD{$p\text{-}D$}
\newcommand\kD{$k\text{-}D$}
\newcommand\dD{$d\text{-}D$}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{agsm}
\usepackage{bookmark}

@@

@@

\begin{itemize}
\tightlist
\item
  NLDR is used to illustrate different ways 1's are drawn
\item
  Use our method to assess is it a reasonable representation
\item
  Demonstrate that it is, except for the anomalies
\end{itemize}

@@

\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pacmapauthor-1.pdf}

@@

\caption{\label{fig-pacmapauthor}(a) 2D layout from PaCMAP applied for
the digit 1 of the MNIST dataset. Is this a best representation of the
original data?, and (b) Images of the handwritten digit 1. The angle of
the digit 1 varies along non-linear structure.}

@@

\caption{\label{fig-model-mnist}(a) Model generated in the 2D space
overlaid on PaCMAP data, and (b) high-D model error in model space.}

@@

@@

\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-anomalies-1.pdf}

@@

\caption{\label{fig-anomalies}(a) Images of handwritten digit 1 which
occur large model error within the non-linear strcuture, and (b) Images
of handwritten digit 1 which occur large model error outside the
non-linear structure.}

>>>>

\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-1in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.7in}%
\addtolength{\topmargin}{-1in}%
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{float}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{bm}
\def\tightlist{}
\usepackage{setspace}
\newcommand\pD{$p\text{-}D$}
\newcommand\kD{$k\text{-}D$}
\newcommand\dD{$d\text{-}D$}
\newcommand\gD{$2\text{-}D$}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{agsm}
\usepackage{bookmark}

@@

\subsubsection{Fitted values}\label{fitted-values}

The prediction approach involves performing the K-nearest neighbors
(KNN) algorithm for an unsupervised classification problem. First, the
nearest high-D model point is identified for a given new high-D point.
Then, the corresponding 2D centroid mapping for the identified high-D
model point is determined. Finally, the coordinates of this 2D centroid
are used as the predicted 2D embedding for the new high-D data point.
This step is particularly valuable due to the limitations of some NLDR
techniques, like tSNE, which don't provide a straightforward method for
prediction. As a result, our approach offers a solution that capable of
generating predicted 2D embedding regardless of the NLDR technique
employed, effectively addressing this functional gap.

\subsubsection{Error calculation}\label{error-calculation}

Residuals are essential for evaluating the accuracy of representing
high-D points by the high-D mapping of 2D bin centroids. To measure this
accuracy, an error metric is introduced, quantifying the sum of squared
differences between the high-D data (\(x_{ij}\)) and the high-D mapping
of the 2D bin centroid data (\(C_{x_ij}\)) across all observations and
dimensions (see Equation~\ref{eq-equation10}).

\begin{equation}\phantomsection\label{eq-equation10}{
\text{Error} = \sum_{j = 1}^{n}\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2
}\end{equation}

Here, \(n\) represents the number of observations, \(p\) represents the
dimensions of high-D data, \(x_{ij}\) is the high-D data, and
\(C_{x_ij}\) is the high-D mapping of the 2D bin centroid.

To assess how well our method captures and represents the underlying
structure of the high-D data, Mean Squared Error (MSE) is used. When
computing MSE, total model error (see \textbf{?@sec-summary}) is divided
by the number of observations to make it as a mean value (see
Equation~\ref{eq-equation9}).

\begin{equation}\phantomsection\label{eq-equation9}{
\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{n}
}\end{equation}

@@

The MNIST dataset consists of grayscale images of handwritten digits
\citep{lecun2010}. \citet{yingfan2021} used this dataset to demonstrate
how PaCMAP preserves local structure. We selected the \gD{} embedding of
PaCMAP for the handwritten digit 1 to assess whether this is a
reasonable representation using our method. As shown in
Figure~\ref{fig-pacmap-author-img}, the angle of the digit 1 images
varies along the \gD{} structure.

@@

\includegraphics[width=1\textwidth,height=0.22\textheight]{paper_files/figure-pdf/fig-pacmap-author-1.pdf}

@@

\caption{\label{fig-pacmap-author}\(2\text{-}D\) layout from PaCMAP
applied for the digit 1 of the MNIST dataset. Is this the best
representation of the digit 1? The parameter setting is n\_components=2,
n\_neighbors=10, init=random, MN\_ratio=0.9, FP\_ratio=2.0.}

\includegraphics[width=1\textwidth,height=0.5\textheight]{paper_files/figure-pdf/fig-pacmap-author-img-1.pdf}

\caption{\label{fig-pacmap-author-img}Images of the handwritten digit 1
are ordered from the bottom-right to the top-left of the \(2\text{-}D\)
structure. The angle of the digit varies along this structure. Images at
the bottom-right of the \(2\text{-}D\) layout show the digit 1 angled
more to the right, while images at the top-left show the digit 1 angled
more to the left. This demonstrates how the angle changes from right to
left along the \(2\text{-}D\) structure.}

@@

\caption{\label{fig-model-mnist}(a) Model generated in \(2\text{-}D\),
and (b) \(p\text{-}D\) model error in \(2\text{-}D\). The \(2\text{-}D\)
model shows a non-linear continuous structure. Most low \(p\text{-}D\)
model errors are distributed along the lower edge of the \(2\text{-}D\)
structure, while most high p-D model errors are concentrated along the
upper edge.}

@@

According to Figure~\ref{fig-mnist1_sc1}, the non-linear continuous
structure observed in the \gD{} representation of PaCMAP (see
Figure~\ref{fig-pacmap-author}) is also visible when visualizing the
model overlaid on the data space. This indicates that PaCMAP accurately
captures the structure of the \pD{} data. Additionally, the model shows
a twisted pattern within the non-linear structure in \pD{} space (see
Figure~\ref{fig-mnist1_sc2}), which is an additional pattern not visible
in the \gD{} representation (see Figure~\ref{fig-pacmap-author}).
Furthermore, as shown in Figure~\ref{fig-mnist1_sc3}, some long edges
exist in the \pD{} space that are not recognized as long edges in the
\gD{} representation. However, PaCMAP is a \emph{good} \gD{}
representation of MNIST digit 1 data.

\begin{minipage}{0.33\linewidth}

\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_1.png}

\subcaption{\label{fig-mnist1_sc1}}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_2.png}

\subcaption{\label{fig-mnist1_sc2}}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_3.png}

\subcaption{\label{fig-mnist1_sc3}}

\end{minipage}%

\caption{\label{fig-mnist1_sc}Screen shots of the \textbf{langevitour}
of the MNIST digit 1 data set, shows the model-in-data space, a video of
the tour animation is available at
(\url{https://youtu.be/zq2GM9qvUNA}).}

There are certain data points that exhibit high error rates due to their
deviation from the usual \pD{} data structure, which makes them
anomalies (see Figure~\ref{fig-model-mnist} (b)). These anomalies can be
classified into two types: those that are anomalies within the
non-linear structure and those that lie outside of it. The images
associated with high model error points within the non-linear structure
display different patterns of the digit 1, as shown in
Figure~\ref{fig-mnist-anomalies} (a). However, when comparing these
images to the ones found outside of the non-linear structure, it becomes
evident that the latter display different patterns of the digit 1 (see
Figure~\ref{fig-mnist-anomalies} (b)).

@@

\includegraphics[width=1\textwidth,height=0.35\textheight]{paper_files/figure-pdf/fig-mnist-anomalies-1.pdf}

@@

\caption{\label{fig-mnist-anomalies}Some images of handwritten digit 1
which occur high model error (a) within the non-linear strcuture, and
(b) outside the non-linear structure. The images shows different
patterns of digit 1.}

### 2024-05-19 supplementary/hexagon_binning_algorithm.qmd
Update with US english
Added MNIST digit description, need to add the video

$$
a_1 > -\frac{4}{\sqrt{3}}qr_2
$$ 
Also, substituting $a_2 = \frac{\sqrt{3}}{2}a_1$, $s_2 = -q \times r_2$ and rearranging @eq-equation5 gives:

@@

Similarly, @eq-equation2 becomes,

@@

This is a linear optimisation problem. Therefore, the optimal solution must occur on a vertex. So, by setting @eq-equation6 equals to @eq-equation7 gives,

@@

```
```{r}
#| label: fig-assign-data
#| fig-cap: "Binning the data. Points are assigned to the nearest centroid. If a point is equidistant from multiple centroids, assigned to the lowest centroid."

>>>>

$$
a_1 > -\frac{4}{\sqrt{3}}qr_2
$$

Also, substituting $a_2 = \frac{\sqrt{3}}{2}a_1$, $s_2 = -q \times r_2$ and rearranging @eq-equation5 gives:

@@

Similarly, substituting $s_1 = -q$ @eq-equation2 becomes,

@@

This is a linear optimization problem. Therefore, the optimal solution must occur on a vertex. So, by setting @eq-equation6 equals to @eq-equation7 gives,

@@

```

```{r}
#| label: fig-assign-data
#| fig-cap: "Binning the data. Points are assigned to the nearest centroid. If a point is equidistant from multiple centroids, assigned to the lowest centroid."
